Smooth risk-based predictive approaches to treatment effect heterogeneity: A simulation study

### 2021-06-09 submission/manuscript.rmd
new set of figures; text not updated

```{r, echo=FALSE, warning=FALSE, message=FALSE}
rmse_scenarios_10_12_16 <- readr::read_csv(
  file = here::here(file.path("data/processed", "rmse.csv"))
) %>%
  dplyr::filter(scenarioId %in% c(10, 12, 16))
```
Under the base case of constant relative treatment effect, the model assuming a
constant treatment effect had the lowest median RMSE, regardless of true
prediction AUC and sample size (Figure \@ref(fig:rmseconstant)). Linear
interaction models demonstrated comparable performance. Among the RCS smoothing
methods, the one fitted with 3 knots always performed best, while the increased
flexibility achieved when increasing the knots resulted in overfitting and worse
performance. The adaptive approach under all scenarios performed similar to the
model with smaller RMSE in all scenarios.

```{r rmseconstant, echo=FALSE, fig.cap="Caption", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/rmse_constant.png"))
```

@@

```{r rmsebase, echo=FALSE, fig.cap="Caption", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/rmse_base.png"))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
rmse_scenarios_10_64_66_65 <- readr::read_csv(
  file = here::here(file.path("data/processed", "rmse.csv"))
) %>%
  dplyr::filter(scenarioId %in% c(10, 64, 66, 65))
```
When focusing on the different scenarios where true treatment-covariate
interactions were considered all methods had similar RMSE performance, (Figure
\@ref(fig:rmseInteractions); Panels B, C and D). In case of strong
treatment-covariate interactions the constant effect model had slightly
increased RMSE
(`r rmse_scenarios_10_64_66_65 %>% filter(scenarioId == 65) %>% select(constant_treatment_effect) %>% unlist() %>% median() %>% round(3)`; [`r rmse_scenarios_10_64_66_65 %>% filter(scenarioId == 65) %>% select(constant_treatment_effect) %>% unlist() %>% quantile(c(.025, .975)) %>% round(3) %>% paste(collapse = ", ")`])
compared to the other methods. The linear interaction model with the risk linear
predictor had the lowest RMSE
(`r rmse_scenarios_10_64_66_65 %>% filter(scenarioId == 65) %>% select(linear_predictor) %>% unlist() %>% median() %>% round(3)`; [`r rmse_scenarios_10_64_66_65 %>% filter(scenarioId == 65) %>% select(linear_predictor) %>% unlist() %>% quantile(c(.025, .975)) %>% round(3) %>% paste(collapse = ", ")`]).

```{r rmseInteractions, echo=FALSE, fig.cap="Root mean squared error of the different methods when focusing on the effect treatment-covariate interactions. In all scenarios sample size is fixed to 4250 and the true prediction AUC is 0.75. \\textit{Panel A} presents the results for the base case scenario, where no interactions are present. \\textit{Panel B} presents the results for the scenario with 4 weak interactions. \\textit{Panel C} presents the results for the scenario with mixed interactions (2 weak and 2 strong). \\textit{Panel D} presents the results for the scenario with 4 strong interactions.", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/rmse_interactions.png"))
```

@@

```{r discrimination, echo=FALSE, fig.cap="Discrimination for benefit of the different methods. \\textit{Panel A} presents the results when moderate linear deviations from the base case scenario of constant treatment effects are considered, sample size is 4250 and true prediction AUC is 0.75. \\textit{Panel B} presents the results when moderate linear deviations are considered and true prediction AUC is 0.85. Sample size is 4250. \\textit{Panel C} presents the results when moderate linear deviations are considered and the sample size is 0.75. \\textit{Panel D} presents the results when strong linear deviations from the base case are considered, sample size is 4250 and true prediction AUC is 0.85. \\textit{Panel E} presents results for true prediction AUC of 0.85 and \\textit{Panel F} for sample size of 17000 (assuming moderate linear deviations from base case). \\textit{Panel G} presents the results when strong quadratic deviations from the base case are considered, sample size is 4250 and true prediction AUC is 0.85. \\textit{Panel H} presents the results for true prediction AUC of 0.85 and \\textit{Panel I} for sample size of 17000 (assuming strong quadratic deviations from base case)", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/discrimination_base.png"))
```

@@

```{r calibration, echo=FALSE, fig.cap="Calibration for benefit of the different methods. \\textit{Panel A} presents the results when moderate linear deviations from the base case scenario of constant treatment effects are considered, sample size is 4250 and true prediction AUC is 0.75. \\textit{Panel B} presents the results when moderate linear deviations are considered and true prediction AUC is 0.85. Sample size is 4250. \\textit{Panel C} presents the results when moderate linear deviations are considered and the sample size is 0.75. \\textit{Panel D} presents the results when strong linear deviations from the base case are considered, sample size is 4250 and true prediction AUC is 0.85. \\textit{Panel E} presents results for true prediction AUC of 0.85 and \\textit{Panel F} for sample size of 17000 (assuming moderate linear deviations from base case). \\textit{Panel G} presents the results when strong quadratic deviations from the base case are considered, sample size is 4250 and true prediction AUC is 0.85. \\textit{Panel H} presents the results for true prediction AUC of 0.85 and \\textit{Panel I} for sample size of 17000 (assuming strong quadratic deviations from base case)", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/calibration_base.png"))
```

>>>>

```{r, echo=FALSE, warning=FALSE, message=FALSE}
rmse_scenarios_10_12_16 <- readr::read_csv(
  file = here::here(file.path("data/processed", "rmse.csv"))
) %>%
  dplyr::filter(scenarioId %in% c(10, 12, 16))
```
Under the base case of constant relative treatment effect, the model assuming a
constant treatment effect had the lowest median RMSE, regardless of true
prediction AUC and sample size. Linear interaction models demonstrated
comparable performance. Among the RCS smoothing methods, the one fitted with 3
knots always performed best, while the increased flexibility achieved when
increasing the knots resulted in overfitting and worse performance. The adaptive
approach under all scenarios performed similar to the model with smaller RMSE in
all scenarios.

```{r rmsebase, echo=FALSE, fig.cap="RMSE base case", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/rmse_base.png"))
```

```{r rmseauc, echo=FALSE, fig.cap="AUC 0.85", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/rmse_auc.png"))
```

```{r rmsesamplesize, echo=FALSE, fig.cap="Sample size 17000", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/rmse_sample_size.png"))
```

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE}
rmse_scenarios_10_64_66_65 <- readr::read_csv(
  file = here::here(file.path("data/processed", "rmse.csv"))
) %>%
  dplyr::filter(scenarioId %in% c(10, 64, 66, 65))
```

When focusing on the different scenarios where true treatment-covariate
interactions were considered all methods had similar RMSE performance. In case
of strong treatment-covariate interactions the constant effect model had
slightly increased RMSE
(`r rmse_scenarios_10_64_66_65 %>% filter(scenarioId == 65) %>% select(constant_treatment_effect) %>% unlist() %>% median() %>% round(3)`; [`r rmse_scenarios_10_64_66_65 %>% filter(scenarioId == 65) %>% select(constant_treatment_effect) %>% unlist() %>% quantile(c(.025, .975)) %>% round(3) %>% paste(collapse = ", ")`])
compared to the other methods. The linear interaction model with the risk linear
predictor had the lowest RMSE
(`r rmse_scenarios_10_64_66_65 %>% filter(scenarioId == 65) %>% select(linear_predictor) %>% unlist() %>% median() %>% round(3)`; [`r rmse_scenarios_10_64_66_65 %>% filter(scenarioId == 65) %>% select(linear_predictor) %>% unlist() %>% quantile(c(.025, .975)) %>% round(3) %>% paste(collapse = ", ")`]).

@@

```{r discrimination, echo=FALSE, fig.cap="Discrimination for benefit base case.", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/discrimination_base.png"))
```

@@

```{r calibration, echo=FALSE, fig.cap="Calibration for benefit base case.", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/calibration_base.png"))
```
